---
title: "Assignment 4"
author: "Brenden Hale"
output: html_notebook
date: "10/27/2022"
---

```{r}
library(tidyverse)  # data manipulation
library(factoextra) # clustering algorithms & visualization
library(ISLR)
library(cluster)
set.seed(123)
```

```{r}
# Load Pharmaceuticals.csv
Pharmaceuticals <- read.csv("Pharmaceuticals.csv")
```

# A. Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 

```{r}
# Use only the numerical variables (3-11)
Pharma <- na.omit(Pharmaceuticals)
df <- Pharma[,c(3:11)]
summary(df)
```

```{r}
# Scaling the data frame's numerical variables
df.scale <- scale(df)
```

```{r}
# In order to find the optimal number of cluters, I will use the Elbow and Silhouette methods. The Elbow is one of the most popular methods but it is inexact. The Silhouette provides a more exact visual of the optimal number of clusters and is fairly straightforward.

# Elbow method to determine clusters
fviz_nbclust(df.scale, kmeans, method = "wss") + labs(subtitle = "Elbow method")

# The eblow appears to pivot at k=4 or k=5
```

```{r}
# Silhouette method to determine clusters
fviz_nbclust(df.scale, kmeans, method = "silhouette")+ labs(subtitle = "Silhouette method")

# According to the Silhouette method 5 is the optimal number of clusters
```

```{r}
# According to the Elbow and Silhouette method shown above, the optimal number of clusters is 5

k5 <- kmeans(df.scale,centers = 5,nstart = 25)
```

```{r}
k5$centers # Centroids for clusters
k5$size
fviz_cluster(k5, data = df.scale) # Visualize clusters
distance <- dist(df.scale, method = "euclidean")
fviz_dist(distance)
```

```{r}
# Visualizations
fit <- kmeans(df.scale,5) # Use K-means to determine fit of the data
aggregate(df.scale, by = list(fit$cluster), FUN=mean) # Calculate mean of all numerical variables in cluster
clusplot(df.scale,fit$cluster, color = TRUE, shade = TRUE, labels = 2, lines = 0)
```

# B. Interpret the clusters with respect to the numerical variables used in forming the clusters.

Based off of the cluster plot above:
Cluster 1- AZN, ABT, NVS, BMY, WYE, SGP, LLY
Cluster 2- GSK, JNJ, MRK, PFE
Cluster 3- CHTT, ELN, MRX, IVX
Cluster 4- AHM, AVE, WPI
Cluster 5- AGN, BAY, PHA

Interpretations based off of data aggregation
Cluster 1 has the highest Net_Profit_Margin and the lowest Rev_Growth, 
Cluster 2 has the highest Market_Cap, ROE, ROA, Asset_Turnover and the lowest Leverage,
Cluster 3 has the highest Beta, Leverage, Rev_Growth and the lowest Asset_Turnover, PE_Ratio, Market_Cap
Cluster 4 has the lowest Beta
Cluster 5 has the highest PE_Ratio and the lowest Net_Profit_Margin, ROA, ROE

# C. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)? (those not used in forming the clusters)
Yes, there is a pattern in regards to the Median_Reccomendation
Cluster 1 is primarily a middle ground of a hold with a few a moderate buy or sell
Cluster 2, 3 and 5 are mostly a moderate buy
Cluster 4 is the most positive buy cluster

# D. Provide an appropriate name for each cluster using any or all of the variables in the dataset.
Cluster 1- High profit low growth
Cluster 2- High market cap
Cluster 3- Smaller market cap high volatility
Cluster 4- Low volatility
Cluster 5- High PE